Declarative Deployment Reconfiguration of Pulverised Collective Adaptive Systems at Runtime \thanks{ This work has been supported by the Italian PRIN Project COMMON-WEARS (2020HCWWLP) and the EU/MUR FSE REACT-EU PON R\&I 2014-2020. }

### 2023-05-21 paper-2023-acsos-pulverization.tex
Minors

@@

In this work,
we further develop the research on pulverised systems by providing two key contributions:
%
\begin{enumerate}
 \item we extend the theoretical framework of pulverisation by adding support for
\emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
 \item we accordigly provide a practical implementation of pulverisation (with reconfiguration rules) called \emph{\ourframework{}}, developed in the Kotlin programming language.
\end{enumerate}
%
To evaluate \ourframework{}, we exercise it through two case studies:
a small-scale real-world demonstration, and a city-scale simulated scenario.

@@

The \emph{deployment view} is a well-known architectural viewpoint for software systems,
concerned with the mapping of \emph{software components} to \emph{physical machines}~\cite{DBLP:journals/software/Kruchten95}
 and supported by modeling notations like \emph{UML Deployment diagrams}.
%
Here, we briefly introduce deployment and reconfiguration based on the conceptual characterisation of~\cite{DBLP:journals/jss/ArcangeliBL15,carzaniga1998characterization}.
%
A \emph{site} is a set of computers (\emph{hosts}) that may host a \emph{software system}, i.e., a product described by a coherent collection of artefacts and consisting of a set of \emph{deployment units} that can be independently operated.
%
\emph{(Software) Deployment} is the process of moving and making a software system available and operational from one or more \emph{producer sites} to a target set of \emph{consumer sites}, also called the \emph{deployment domain}.
%
The \emph{deployment plan} defines 
 the mapping between the software system
 and the deployment domain,
 possibly augmented with further information (e.g., metadata, constraints, preferences).
%
Common deployment-related activities include:
(i) \emph{release/update} of the software system at the producer site;
(ii) \emph{installation/uninstallation} at/from the consumer sites;
(iii) \emph{activation/deactivation}, for starting/stopping the components;
(iv) \emph{reorganisation} of the software system;
and (v) \emph{redistribution}, i.e., changing the deployment plan.
%
According to the analytical framework of~\cite{DBLP:journals/jss/ArcangeliBL15},
  the problem of (automatic) deployment of distributed software systems can be addressed by considering
 (i) the nature of the software to be deployed (e.g., how the software system is split into components);
 (ii) the nature of the deployment domain (i.e. the characteristics and topology of the available infrastructure);
 (iii) how the deployment is designed (e.g., how the deployment plan is specified);
 and 
 (iv) how the deployment is performed (e.g., how deployment activities are carried out).
%
From an operational point of view,
 deployment may be supported by a so-called \emph{runtime} or \emph{middleware}~\cite{DBLP:journals/cacm/GazisK22}.
%
Related work covering these issues is provided in \Cref{sec:rw}.
%
In the following, we introduce the deployment approach 
  that we extend and upon which we develop \ourframework{}.

@@

We implemented the proposed \ac{dsl} in Kotlin.
%
We wanted a language with a strongly typed syntax to help developers with a better contextual assist,
and we wanted the possibility to target multiple runtimes
(Kotlin currently targets the JVM, JavaScript, and native code for many platforms\footnote{\url{https://archive.is/JSiTg}},
including mobile and wearable devices)
and be as interoperable as possible with other languages.
%
Scala could have been a good alternative,
but we deemed the multiplatform support of Kotlin a better fit for our needs.
%
The framework has been open-sourced and released with a permissive license\footnote{\url{https://github.com/pulvreakt/pulvreakt}}.

@@

In this section,
we exercise the proposed framework through two case studies.
%
In the former,
we show a small-scale demonstrator of a safety system warning people when a room capacity is exceeded
to showcase the integration of the framework with established technologies\footnote{
    \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-crowd-room}
};
in the latter,
we simulate a city-scale distributed application deployed for an urban event
and show how the dynamic reconfiguration may help to balance
cost, energy consumption and \ac{qos}\footnote{
    \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-evaluation}
}.
%
Both experiments were released with a permissive open-source license and archived for future reference on
Zenodo~\cite{https://doi.org/10.5281/zenodo.7933160, https://doi.org/10.5281/zenodo.7948801}.

@@

\subsection{Large-Scale Scenario: Collective Computation at an Urban Event}

@@

\subsubsection{Results}
%
\begin{figure}
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=100.0.pdf}\\
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=300.0.pdf}\\
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=1000.0.pdf}
    \caption{
        Simulation results for 100 (top), 300 (center), and 1000 (bottom) devices;
        showing all four metrics:
        $P_{cloud}$ ($kW$) in green,
        $P_{devices}$ ($kW$) in blue,
        Distance in yellow, $\$_{cloud}$ in dark purple.
        %
        $P_{cloud}$ and $P_{devices}$ are stacked to show the overall power consumption.
        %
        Different reconfiguration rules are indicated on the horizontal axis.
        %
        Runtime reconfiguration of the deployment opens more opportunities to find the
        cost, energy use, and \ac{qos} trade-off that best fits the current context.
    }
    \label{fig:results}
\end{figure}
%
Results are summarised in \Cref{fig:results}.
%
$\Downarrow$ has the worst \ac{qos}
but also the lowest cloud cost and overall energy consumption.
%
Opposedly, $\Uparrow$ improves the \ac{qos}, but costs and energy use grow significantly.
%
The ability to reconfigure the deployment at runtime
by moving the behaviour from the cloud to the smartphones and vice versa
depending on the battery level allows us to find different trade-offs between these two extremes.

@@

A related \ac{dsl} is 
\emph{MuScADeL}~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}:
it allows to express deployment properties of
applications on multi-scale deployment domains,
i.e., large and heterogeneous infrastructures considered under several viewpoints
(device, geography, network, administrative domains, etc.).
%
Component descriptions can include \emph{constraints}
(both basic software/hardware constraints and multi-scale criteria).
%
The specifications are used to generate \emph{probe} artefacts supporting data collection at the level of the deployment middleware.
%
However, the work in~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel} is preliminary and does not cover runtime adaptation.

@@

In its current form, the proposed \ac{dsl} has two limitations that will be addressed in the future.
%
First and foremost, it currently does not consider openness,
as hosts get specified in the runtime and reconfiguration \acp{dsl} as a closed set.
%
The second limitation is related to the policies that can be expressed in the reconfiguration \ac{dsl}:
currently, only policies that can be assessed at the local host level can be assessed.
%
A policy that requires to consider the state of multiple hosts would require to resort to the host language's primitive mechanisms and make explicit communications outside the framework,
thus breaking the abstraction.
%
This limitation can be addressed by extending the current reconfiguration
block of the \ac{dsl} so that queries can be performed on any reachable host of the system,
thus making all communications
(including reconfiguration-related ones)
pass through the channels controlled by the framework.
%
Concurrently with these extensions,
we intend to add support for additional network protocols to support a larger number of practical scenarios.

>>>>

%\usepackage[none]{hyphenat}

@@

In this work,
we further develop the research on pulverised systems by providing two key contributions:
%
\begin{enumerate}
 \item we extend the theoretical framework of pulverisation by adding support for
\emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
 \item we accordingly provide a practical implementation of pulverisation (with reconfiguration rules) called \emph{\ourframework{}}, developed in the Kotlin programming language.
\end{enumerate}
%
To evaluate \ourframework{}, we exercise it through two case studies:
a small-scale real-world demonstration, and a city-scale simulated scenario.

@@

The \emph{deployment view} is a well-known architectural viewpoint for software systems,
concerned with the mapping of \emph{software components} to \emph{physical machines}~\cite{DBLP:journals/software/Kruchten95}
 and supported by modelling notations like \emph{UML Deployment diagrams}.
%
Here, we briefly introduce deployment and reconfiguration based on the conceptual characterisation of~\cite{DBLP:journals/jss/ArcangeliBL15,carzaniga1998characterization}.
%
A \emph{site} is a set of computers (\emph{hosts}) that may host a \emph{software system}, i.e., a product described by a coherent collection of artefacts and consisting of a set of \emph{deployment units} that can be independently operated.
%
\emph{(Software) Deployment} is the process of moving and making a software system available and operational from one or more \emph{producer sites} to a target set of \emph{consumer sites}, also called the \emph{deployment domain}.
%
The \emph{deployment plan} defines 
 the mapping between the software system
 and the deployment domain,
 possibly augmented with further information (e.g., metadata, constraints, preferences).
%
Common deployment-related activities include:
(i) \emph{release/update} of the software system at the producer site;
(ii) \emph{installation/uninstallation} at/from the consumer sites;
(iii) \emph{activation/deactivation}, for starting/stopping the components;
(iv) \emph{reorganisation} of the software system;
and (v) \emph{redistribution}, i.e., changing the deployment plan.
%
According to the analytical framework of~\cite{DBLP:journals/jss/ArcangeliBL15},
  the problem of (automatic) deployment of distributed software systems can be addressed by considering
 (i) the nature of the software to be deployed (e.g., how the software system is split into components);
 (ii) the nature of the deployment domain (i.e. the characteristics and topology of the available infrastructure);
 (iii) how the deployment is designed (e.g., how the deployment plan is specified);
 and 
 (iv) how the deployment is performed (e.g., how deployment activities are carried out).
%
From an operational point of view,
 deployment may be supported by a so-called \emph{runtime} or \emph{middleware}~\cite{DBLP:journals/cacm/GazisK22}.
%
Related work covering these issues is provided in \Cref{sec:rw}.
%
In the following, we introduce the deployment approach 
  that we extend and upon which we develop \ourframework{}.

@@

We implemented the proposed \ac{dsl} in Kotlin.
%
We wanted a language with a strongly typed syntax to help developers with a better contextual assist,
and we wanted the possibility to target multiple runtimes
(Kotlin currently targets the JVM, JavaScript, and native code for many platforms\footnote{\url{https://archive.is/JSiTg}},
including mobile and wearable devices)
and be as interoperable as possible with other languages.
%
Scala could have been a good alternative,
but we deemed the multiplatform support of Kotlin a better fit for our needs.
%
The framework has been open-sourced and released with a permissive licence\footnote{\url{https://github.com/pulvreakt/pulvreakt}}.

@@

In this section,
we exercise the proposed framework through two case studies.
%
In the former,
we show a small-scale demonstrator of a safety system warning people when a room capacity is exceeded
to showcase the integration of the framework with established technologies\footnote{
    \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-crowd-room}
};
in the latter,
we simulate a city-scale distributed application deployed for a urban event
and show how the dynamic reconfiguration may help to balance
cost, energy consumption and \ac{qos}\footnote{
    \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-evaluation}
}.
%
Both experiments were released with a permissive open-source licence and archived for future reference on
Zenodo~\cite{https://doi.org/10.5281/zenodo.7933160, https://doi.org/10.5281/zenodo.7948801}.

@@

\subsection{Large-Scale Scenario: Collective Computation at a Urban Event}

@@

\subsubsection{Results}
%
\begin{figure}
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=100.0.pdf}\\
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=300.0.pdf}\\
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=1000.0.pdf}
    \caption{
        Simulation results for 100 (top), 300 (center), and 1000 (bottom) devices;
        showing all four metrics:
        $P_{cloud}$ ($kW$) in green,
        $P_{devices}$ ($kW$) in blue,
        Distance in yellow, $\$_{cloud}$ in dark purple.
        %
        $P_{cloud}$ and $P_{devices}$ are stacked to show the overall power consumption.
        %
        Different reconfiguration rules are indicated on the horizontal axis.
        %
        Runtime reconfiguration of the deployment opens more opportunities to find the
        cost, energy use, and \ac{qos} trade-off that best fits the current context.
    }
    \label{fig:results}
\end{figure}
%
Results are summarised in \Cref{fig:results}.
%
$\Downarrow$ has the worst \ac{qos}
but also the lowest cloud cost and overall energy consumption.
%
Conversely, $\Uparrow$ improves the \ac{qos}, but costs and energy use grow significantly.
%
The ability to reconfigure the deployment at runtime
by moving the behaviour from the cloud to the smartphones and vice versa
depending on the battery level allows us to find different trade-offs between these two extremes.

@@

A related \ac{dsl} is 
\emph{MuScADeL}~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}:
it allows expressing deployment properties of
applications on multi-scale deployment domains,
i.e., large and heterogeneous infrastructures considered under several viewpoints
(device, geography, network, administrative domains, etc.).
%
Component descriptions can include \emph{constraints}
(both basic software/hardware constraints and multi-scale criteria).
%
The specifications are used to generate \emph{probe} artefacts supporting data collection at the level of the deployment middleware.
%
However, the work in~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel} is preliminary and does not cover runtime adaptation.

@@

In its current form, the proposed \ac{dsl} has two limitations that will be addressed in the future.
%
First and foremost, it currently does not consider openness,
as hosts get specified in the runtime and reconfiguration \acp{dsl} as a closed set.
%
The second limitation is related to the policies that can be expressed in the reconfiguration \ac{dsl}:
currently, only policies that can be assessed at the local host level can be assessed.
%
A policy that requires to consider the state of multiple hosts would require resorting to the host language's primitive mechanisms and make explicit communications outside the framework,
thus breaking the abstraction.
%
This limitation can be addressed by extending the current reconfiguration
block of the \ac{dsl} so that queries can be performed on any reachable host of the system,
thus making all communications
(including reconfiguration-related ones)
pass through the channels controlled by the framework.
%
Concurrently with these extensions,
we intend to add support for additional network protocols to support a larger number of practical scenarios.
